`flippant` --- An `R` package for the automated analysis of dithionite scramblase

### 2016-06-16 inst/manuscript/manuscript.rmd
reformat "R" and package names;

---
title: '*flippant* --- An R package for the automated analysis of dithionite scramblase
  assays'
author: Richard J. Cotton ^†^, Birgit Ploier ^‡^, Michael A. Goren ^‡^,
  Anant K. Menon ^‡^ and Johannes Graumann ^†,^*
date: '`r format(Sys.Date(), "%B %d,%Y")`'
output:
  pdf_document:
    fig_caption: yes
  word_document:
    fig_caption: yes
  html_document:
    fig_caption: yes
csl: bioinformatics.csl
bibliography: flippant.bib
---

@@

```{r setting_up_r, echo = FALSE}
library(dplyr, quietly = TRUE)
library(flippant, quietly = TRUE)
library(grid, quietly = TRUE)
library(magrittr, quietly = TRUE)
library(microbenchmark, quietly = TRUE)
library(multipanelfigure, quietly = TRUE)
library(stringi, quietly = TRUE)
flippant_dependencies <- packageDescription("flippant",fields=c("Depends","Imports")) %>%
  unlist(use.names = FALSE) %>%
  stri_split(regex = ",", omit_empty = TRUE) %>%
  unlist(use.names = FALSE) %>%
  stri_replace_all(replacement = "", regex = "\\s*\\(.*\\)\\s*", omit_empty = TRUE) %>%
  stri_replace_all(replacement = "", regex = "\\s*", omit_empty = TRUE) %>%
  tail(n=length(.) - 1) %>%
  sort() %>%
  ( function(x){
    paste0(
      c(
        "*",
        paste0(head(x, n = length(x) - 1), collapse = "*, *"),
        "* and *",
        tail(x, n = 1),
        "*"),
      collapse = "")} )

@@

# Abstract
Dithionite scramblase assays are an established means of probing the lipid 
scrambling activity of protein extracts and purified proteins. While the assay
has yielded insight into the scramblase activity in crude membrane preparations,
functional validation of candidate scramblases, stoichiometry of scramblase 
complexes and probe ATP--dependence of flippases, data analysis in its context 
has remained a task involving many manual steps. With the extension package 
*flippant* to R, a free software environment for statistical computing and 
graphics, we introduce an integrated solution for the analysis and 
publication--grade graphical presentation of dithionite scramblase assays and 
demonstrate its utility in revisiting an originally manual analysis from the 
publication record, closely reproducing the reported results.

@@

Starting with the identification of baseline and post--reduction fluorescence 
from fluorometer--generated spectra, via the calculation of the underlying 
statistics and through publication--grade representation of the results, data 
analysis in the context of this assay has until now involved an extensive series
of manual steps, at most supported by spreadsheet calculation facilities
[see @menon_opsin_2011;@goren_constitutive_2014 for extensive step--wise discussion].
Based on the R free software environment for statistical computing and 
graphics [@r_core_team_r:_2015] and prominently drawing on the power and 
elegance of the `ggplot2` plotting system [@wickham_ggplot2_2009], we here 
provide *flippant*, a package that comprehensively, fast and reproducibly covers
all data analysis and graphing needs from raw fluorescence spectra produced by 
scramblase assays. A simple parametrizing spreadsheet is the sole required user 
input.

# Availablility and Implementation
*flippant* has been implemented in R [@r_core_team_r:_2015] and integrated 
into the `bioconductor` collection of tools for biomedical data analysis [@huber_orchestrating_2015]. It is thus install-- and loadable as follows from 
within a functional R environment:

@@

The package depends on the following packaged extensions to R base 
functionality, which are automatically satisfied on installation: 
`r flippant_dependencies`.

# Input Data and Data Processing
To analyse raw data stemming from dithionite scramblase assays, *flippant* 
requires for each data point information on the path to the 
fluorometer--generated spectrum, the amount of protein used to reconstitute 
proteoliposomes, the assay volume before and after addition of 
dithionite (to calculate a volume correction factor), the amount of lipids 
present (to calculate a protein to phospholipid ratio or PPR) etc. The user 
supplies this information in a spreadsheet (see below for supporting 
functionality).

Spectral input data is read by *flippant* using the file paths provided in the
spreadsheet. Raw spectra produced by QuantaMaster fluorometer instruments 
(Photon Technology International, Inc., Edison, New Jersey) running software 
versions FelixGX v4.1 and Felix32 v1.20, respectively, as well as a simple 
generic tab--delimited spectrum representing file format with time (in seconds) 
and fluorescence intensity representing columns are currently supported. The 
latter is trivially assembled from any fluorescence data using a spreadsheet 
program. File type determination is handled algorithmically.

Based on @menon_opsin_2011, @goren_constitutive_2014 and
@ploier_perturbed_submitted and largely keeping with nomenclature used therein,
data is processed as follows. Input is format checked and defaults are injected
for facultative parameters/columns as appropriate. 
Fluorescence spectra are parsed. This includes algorithmic determination of when 
dithionite was added to the sample using peak detection by continuous wavelet
transformation as implemented in the *wmtsa* R package [@percival_wavelet_2000;@constantine_wmtsa:_2013]. 
Acquisition time is aligned for all spectra in the data set such that the zero 
time point henceforth corresponds to the time of addition. 
Pre--dithionite--addition baseline fluorescence $F_{\text{baseline}}$ is 
determined for each spectrum by averaging (median) over the ten values preceding
dithionite addition. Post--dithionite--addition minimum fluorescence 
$F_{\text{min}}$ is analogously calculated from the last ten data points before 
(and including) a default 400 s or the time point of measurement supplied by the
user in the parameter spreadsheet. $F_{\text{min}}$ is volume--corrected based on
supplied reaction volumes with and without dithionite and for each spectrum/data
point a fluorescence reduction is calculated as follows:
$$
y_{\text{raw}} = 1 - \frac{F_{\text{min}}}{F_{\text{baseline}}}
$$
A relative fluorescence reduction $y$ is calculated scaled to a (required) liposomes--only/no--protein control and the probability $P_{\geq 1}$ for a 
liposome to contain at least one scramblase molecules is calculated using
$$
P_{\geq 1} = \frac{(y - y_0)}{(y_{\text{max}} - y_0)}
$$
where $y$ is the relative fluorescence reduction for the proteoliposome 
measurement and $y_0$ for a measurement of protein--free liposomes. 
$y_{\text{max}}$ refers to either the maximal $y$ in the experiment or is 
derived from a monoexponential fit to the data (default) as a precaution for the 
case where the protein/phospholipid titration did not reach the plateau of the 
saturation curve. The behavior is adjustable using the `scale_to` parameter of 
the corresponding functions (see below). Next a monoexponential curve is fitted
to the relationship of $P_{\geq 1}$ to the protein to phospholipid ratio (PPR; 
(mg/mmol)), providing the scramblase--characterizing fit factor $\alpha$. By
default and following @ploier_perturbed_submitted, PPR is scaled by a factor of
0.65 to account for a fraction of the vesicle pool used that is refractory to
scramblase reconstitution. This behavior can be modified using the 
`ppr_scale_factor` parameter and avoided alltogether setting it to `NULL`.

@@

The application agnostic tab--delimited format must be maintained for the 
completed table, which is subjected to extensive input checks when read by 
*flippant*.

@@

and handing the function the path to the user assembled spreadsheet of input 
data (`x`). The `adjust` argument to the function call serves to selectively 
inactivate *wmtsa*--provided (see above) alignment of multiple spectra to the 
time of dithionite addition, while the `ppr_scale_factor` parameter is used to
scale PPR (see above) and is needed here as the traces are colored by that
measure. Acquisition time coverage of the plot may be tuned  by user--provided
minimal (`time_min_sec` ) and maximal (`time_max_sec`) time points to be
included (in seconds).

@@

*flippant*s capabilities are demonstrated here by reanalyzing data underlying 
Figure 4C/D in @ploier_perturbed_submitted. The following work flow in a first 
step plots the corresponding spectral traces (Figure 1A). Note that algorithmic
alignment to the time of dithionite addition is employed (default) as well as 
custom trimming of the time axis to account for divergent acquisition 
times between experiments. PPR scaling is used with the default factor of 0.65.

@@

```{r figure1_creation, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "***flippant*--based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblase_assay_traces`, while (B) and (C) show protein to phospholipid ration (PPR) plots for the data using a x--axis trimmed to a maximum value of 2 (see text for details). In (B) data is separately plotted for independent experiments, while (C) combines all data points, aiming for reliability of fit. 'adj. PPR' indicates that the measure has been scaled to account for a vesicle pool refractory to flippase reconsitution."}
x <- tmpDir %>%
  file.path("inputTable.txt") %>%
  flippant:::read_scramblase_input_file()
x[["Path"]] <- file.path(tmpDir, x[["Path"]])
figure1 <- add_panel(
  figure = figure1,
  top_panel = 1,
  bottom_panel = 3,
  left_panel = 1, 
  right_panel = 3, 
  panel = scramblase_assay_traces(x, time_max_sec = 350) + theme_bw())
figure1 <- add_panel(
  figure = figure1,
  top_panel = 4,
  bottom_panel = 5,
  left_panel = 1, 
  right_panel = 3, 
  panel = scramblase_assay_plot(x, generation_of_algorithm = 2) + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
figure1 <- add_panel(
  figure = figure1,
  top_panel = 6,
  bottom_panel = 7,
  left_panel = 1,
  right_panel = 3,
  panel = scramblase_assay_plot(x, generation_of_algorithm = 2, split_by_experiment = FALSE) + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])) +
    facet_wrap(c("`Experimental Series`"))+
    guides(colour = "none"))
grid.draw(figure1)
```

@@

```{r analysis_tabular_output, echo = FALSE}
ppr_results <- scramblase_assay_stats(
    x, generation_of_algorithm = 2, split_by_experiment = FALSE) %>%
  mutate(`Fit Constant` = round(`Fit Constant` * 10^4, 2)) %>%
  set_names(c("Experimental Series", "Fit Constant ($\\times 10^4$)"))
ppr_results_names <- ppr_results %>%
  extract2("Experimental Series") %>%
  head(n = nrow(ppr_results) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results$`Experimental Series`, n = 1))
ppr_results_alpha_print <- ppr_results %>%
  extract2("Fit Constant ($\\times 10^4$)") %>%
  round(2) %>%
  head(n = nrow(ppr_results) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results$`Fit Constant ($\\times 10^4$)`, n = 1))
knitr::kable(
  ppr_results,
  row.names = FALSE,
  caption = "**Tabular output of *flippant*--based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** An plain text--version of this output is produced by calling `scramblase_assay_stats`.")
```

The resulting values for the fit constant $\alpha$ correspond with
`r ppr_results_alpha_print` for `r ppr_results_names`, respectively, well to the
prior manual analysis by @ploier_perturbed_submitted (Table S2), where $\alpha
\times 10^4 \approx 16$ was reported for mutant and $\alpha \times 10^4 \approx
7$ for wild type rhodopsin. Based on the manual analysis of the data we conclude
in @ploier_perturbed_submitted that the mutant rhodopsins investigated are
reconstituted into proteoliposomes in monomeric form, contrary to the wild type,
which is reconstituting as a homodimer. 
*flippant* analysis of the data shows identical trends and close matches of the
results, supporting identical conclusions.

@@

With *flippant* we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
R as an environment for statistical computing and graphics 
[@r_core_team_r:_2015], scientists can quickly analyse such data and arrive at 
publication--grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
R version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr`~\text{s}$; $n=`r analysis_timing_n`$).

Results from *flippant*--driven analysis track prior manual analysis well. 
Deviations observed may stem from algorithmic differences such as the fitting 
routine used (*flippant* uses the the Marquardt nonlinear least squares
algorithm as implemented by @elzhov_minpack.lm:_2015) and 
general handling of significant digits and rounding.

The choice of R for the implementation is not only motivated by the extensive
tool kit for statistic methodology and exquisite plotting environment it 
provides, but also by the fact that it is a scripting language, rendering the
implementation unequivocally open to scrutiny and adaptation or extension by 
users.

*flippant*, in conclusion, provides means for quick, reproducible data analysis
in the context of dithionite scramblase analysis and a platform for review, 
dissemination and extension of the strategies it employs. As BSA--back 
extraction, a different assay for lipid translocation across liposome bilayers
[@kubelt_transbilayer_2002;@vehring_flip-flop_2007], can utilize an essentially 
identical data analysis workflow, *flippant* is expected to facilitate analysis
of such experiments as well.

>>>>

---
title: '`flippant` --- An `R` package for the automated analysis of dithionite scramblase
  assays'
author: Richard J. Cotton ^†^, Birgit Ploier ^‡^, Michael A. Goren ^‡^,
  Anant K. Menon ^‡^ and Johannes Graumann ^†,^*
date: '`r format(Sys.Date(), "%B %d,%Y")`'
output:
  pdf_document:
    fig_caption: yes
  word_document:
    fig_caption: yes
  html_document:
    fig_caption: yes
csl: bioinformatics.csl
bibliography: flippant.bib
---

@@

```{r setting_up_r, echo = FALSE}
library(dplyr, quietly = TRUE)
library(flippant, quietly = TRUE)
library(grid, quietly = TRUE)
library(magrittr, quietly = TRUE)
library(microbenchmark, quietly = TRUE)
library(multipanelfigure, quietly = TRUE)
library(stringi, quietly = TRUE)
flippant_dependencies <- packageDescription("flippant",fields=c("Depends","Imports")) %>%
  unlist(use.names = FALSE) %>%
  stri_split(regex = ",", omit_empty = TRUE) %>%
  unlist(use.names = FALSE) %>%
  stri_replace_all(replacement = "", regex = "\\s*\\(.*\\)\\s*", omit_empty = TRUE) %>%
  stri_replace_all(replacement = "", regex = "\\s*", omit_empty = TRUE) %>%
  tail(n=length(.) - 1) %>%
  sort() %>%
  ( function(x){
    paste0(
      c(
        "`",
        paste0(head(x, n = length(x) - 1), collapse = "`, `"),
        "` and `",
        tail(x, n = 1),
        "`"),
      collapse = "")} )

@@

# Abstract
Dithionite scramblase assays are an established means of probing the lipid 
scrambling activity of protein extracts and purified proteins. While the assay
has yielded insight into the scramblase activity in crude membrane preparations,
functional validation of candidate scramblases, stoichiometry of scramblase 
complexes and probe ATP--dependence of flippases, data analysis in its context 
has remained a task involving many manual steps. With the extension package 
`flippant` to `R`, a free software environment for statistical computing and 
graphics, we introduce an integrated solution for the analysis and 
publication--grade graphical presentation of dithionite scramblase assays and 
demonstrate its utility in revisiting an originally manual analysis from the 
publication record, closely reproducing the reported results.

@@

Starting with the identification of baseline and post--reduction fluorescence 
from fluorometer--generated spectra, via the calculation of the underlying 
statistics and through publication--grade representation of the results, data 
analysis in the context of this assay has until now involved an extensive series
of manual steps, at most supported by spreadsheet calculation facilities
[see @menon_opsin_2011;@goren_constitutive_2014 for extensive step--wise discussion].
Based on the `R` free software environment for statistical computing and 
graphics [@r_core_team_r:_2015] and prominently drawing on the power and 
elegance of the `ggplot2` plotting system [@wickham_ggplot2_2009], we here 
provide `flippant`, a package that comprehensively, fast and reproducibly covers
all data analysis and graphing needs from raw fluorescence spectra produced by 
scramblase assays. A simple parametrizing spreadsheet is the sole required user 
input.

# Availablility and Implementation
`flippant` has been implemented in `R` [@r_core_team_r:_2015] and integrated 
into the `bioconductor` collection of tools for biomedical data analysis [@huber_orchestrating_2015]. It is thus install-- and loadable as follows from 
within a functional `R` environment:

@@

The package depends on the following packaged extensions to `R` base 
functionality, which are automatically satisfied on installation: 
`r flippant_dependencies`.

# Input Data and Data Processing
To analyse raw data stemming from dithionite scramblase assays, `flippant` 
requires for each data point information on the path to the 
fluorometer--generated spectrum, the amount of protein used to reconstitute 
proteoliposomes, the assay volume before and after addition of 
dithionite (to calculate a volume correction factor), the amount of lipids 
present (to calculate a protein to phospholipid ratio or PPR) etc. The user 
supplies this information in a spreadsheet (see below for supporting 
functionality).

Spectral input data is read by `flippant` using the file paths provided in the
spreadsheet. Raw spectra produced by QuantaMaster fluorometer instruments 
(Photon Technology International, Inc., Edison, New Jersey) running software 
versions FelixGX v4.1 and Felix32 v1.20, respectively, as well as a simple 
generic tab--delimited spectrum representing file format with time (in seconds) 
and fluorescence intensity representing columns are currently supported. The 
latter is trivially assembled from any fluorescence data using a spreadsheet 
program. File type determination is handled algorithmically.

Based on @menon_opsin_2011, @goren_constitutive_2014 and
@ploier_perturbed_submitted and largely keeping with nomenclature used therein,
data is processed as follows. Input is format checked and defaults are injected
for facultative parameters/columns as appropriate. 
Fluorescence spectra are parsed. This includes algorithmic determination of when 
dithionite was added to the sample using peak detection by continuous wavelet
transformation as implemented in the `wmtsa` package [@percival_wavelet_2000;@constantine_wmtsa:_2013]. 
Acquisition time is aligned for all spectra in the data set such that the zero 
time point henceforth corresponds to the time of addition. 
Pre--dithionite--addition baseline fluorescence $F_{\text{baseline}}$ is 
determined for each spectrum by averaging (median) over the ten values preceding
dithionite addition. Post--dithionite--addition minimum fluorescence 
$F_{\text{min}}$ is analogously calculated from the last ten data points before 
(and including) a default 400 s or the time point of measurement supplied by the
user in the parameter spreadsheet. $F_{\text{min}}$ is volume--corrected based on
supplied reaction volumes with and without dithionite and for each spectrum/data
point a fluorescence reduction is calculated as follows:
$$
y_{\text{raw}} = 1 - \frac{F_{\text{min}}}{F_{\text{baseline}}}
$$
A relative fluorescence reduction $y$ is calculated scaled to a (required) liposomes--only/no--protein control and the probability $P_{\geq 1}$ for a 
liposome to contain at least one scramblase molecules is calculated using
$$
P_{\geq 1} = \frac{(y - y_0)}{(y_{\text{max}} - y_0)}
$$
where $y$ is the relative fluorescence reduction for the proteoliposome 
measurement and $y_0$ for a measurement of protein--free liposomes. 
$y_{\text{max}}$ refers to either the maximal $y$ in the experiment or is 
derived from a monoexponential fit to the data (default) as a precaution for the 
case where the protein/phospholipid titration did not reach the plateau of the 
saturation curve. The behavior is adjustable using the `scale_to` parameter of 
the corresponding functions (see below). Next a monoexponential curve is fitted
to the relationship of $P_{\geq 1}$ to the protein to phospholipid ratio (PPR; 
(mg/mmol)), providing the scramblase--characterizing fit factor $\alpha$. By
default and following @ploier_perturbed_submitted, PPR is scaled by a factor of
0.65 to account for a fraction of the vesicle pool used that is refractory to
scramblase reconstitution. This behavior can be modified using the 
`ppr_scale_factor` parameter and avoided alltogether setting it to `NULL`.

@@

The application agnostic tab--delimited format must be maintained for the 
completed table, which is subjected to extensive input checks when read by 
`flippant`.

@@

and handing the function the path to the user assembled spreadsheet of input 
data (`x`). The `adjust` argument to the function call serves to selectively 
inactivate `wmtsa`--provided (see above) alignment of multiple spectra to the 
time of dithionite addition, while the `ppr_scale_factor` parameter is used to
scale PPR (see above) and is needed here as the traces are colored by that
measure. Acquisition time coverage of the plot may be tuned  by user--provided
minimal (`time_min_sec` ) and maximal (`time_max_sec`) time points to be
included (in seconds).

@@

Figure 4C/D in @ploier_perturbed_submitted. The following work flow in a first 
`flippant`s capabilities are demonstrated here by reanalyzing data underlying 
Figure 4C--F in @ploier_perturbed_submitted. The following work flow in a first 
alignment to the time of dithionite addition is employed (default) as well as 
custom trimming of the time axis to account for divergent acquisition 
times between experiments. PPR scaling is used with the default factor of 0.65.

@@

```{r figure1_creation, echo = FALSE, fig.height = 4, fig.cap = "**`flippant`--based plotting of spectral traces underlying the data in Figure 4C--F from @ploier_perturbed_submitted.** See text for command and details."}
# Plot spectral traces (Figure 1)
x <- tmpDir %>%
  file.path("inputTable.txt") %>%
  flippant:::read_scramblase_input_file()
x[["Path"]] <- file.path(tmpDir, x[["Path"]])
figure1 <- add_panel(
  figure = figure1,
  top_panel = 1,
  bottom_panel = 3,
  left_panel = 1, 
  right_panel = 3, 
  panel = scramblase_assay_traces(x, time_max_sec = 350) + theme_bw())
figure1 <- add_panel(
  figure = figure1,
  top_panel = 4,
  bottom_panel = 5,
  left_panel = 1, 
  right_panel = 3, 
  panel = scramblase_assay_plot(x, generation_of_algorithm = 2) + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
figure1 <- add_panel(
  figure = figure1,
  top_panel = 6,
  bottom_panel = 7,
  left_panel = 1,
  right_panel = 3,
  panel = scramblase_assay_plot(x, generation_of_algorithm = 2, split_by_experiment = FALSE) + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("adj. PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])) +
    facet_wrap(c("`Experimental Series`"))+
    guides(colour = "none"))
grid.draw(figure1)
```

@@

```{r analysis_tabular_output, echo = FALSE}
ppr_results <- scramblase_assay_stats(
    x, generation_of_algorithm = 2, split_by_experiment = FALSE) %>%
  mutate(`Fit Constant` = round(`Fit Constant` * 10^4, 2)) %>%
  set_names(c("Experimental Series", "Fit Constant ($\\times 10^4$)"))
ppr_results_names <- ppr_results %>%
  extract2("Experimental Series") %>%
  head(n = nrow(ppr_results) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results$`Experimental Series`, n = 1))
ppr_results_alpha_print <- ppr_results %>%
  extract2("Fit Constant ($\\times 10^4$)") %>%
  round(2) %>%
  head(n = nrow(ppr_results) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results$`Fit Constant ($\\times 10^4$)`, n = 1))
knitr::kable(
  ppr_results,
  row.names = FALSE,
  caption = "**Tabular output of `flippant`--based reanalysis of data in Figure 4C--F from @ploier_perturbed_submitted.** An plain text--version of this output is produced by calling `scramblase_assay_stats`.")
```

The resulting values for the fit constant $\alpha$ correspond with
`r ppr_results_alpha_print` for `r ppr_results_names`, respectively, well to the
prior manual analysis by @ploier_perturbed_submitted (Table S2), where $\alpha
\times 10^4 \approx 16$ was reported for mutant and $\alpha \times 10^4 \approx
7$ for wild type rhodopsin. Based on the manual analysis of the data we conclude
in @ploier_perturbed_submitted that the mutant rhodopsins investigated are
reconstituted into proteoliposomes in monomeric form, contrary to the wild type,
which is reconstituting as a homodimer. 
`flippant` analysis of the data shows identical trends and close matches of the
results, supporting identical conclusions.

@@

With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
R as an environment for statistical computing and graphics 
[@r_core_team_r:_2015], scientists can quickly analyse such data and arrive at 
publication--grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
`R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr`~\text{s}$; $n=`r analysis_timing_n`$).

Results from `flippant`--driven analysis track prior manual analysis well. 
Deviations observed may stem from algorithmic differences such as the fitting 
routine used (`flippant` uses the the Marquardt nonlinear least squares
algorithm as implemented by @elzhov_minpack.lm:_2015) and 
general handling of significant digits and rounding.

The choice of `R` for the implementation is not only motivated by the extensive
tool kit for statistic methodology and exquisite plotting environment it 
provides, but also by the fact that it is a scripting language, rendering the
implementation unequivocally open to scrutiny and adaptation or extension by 
users.

`flippant`, in conclusion, provides means for quick, reproducible data analysis
in the context of dithionite scramblase analysis and a platform for review, 
dissemination and extension of the strategies it employs. As BSA--back 
extraction, a different assay for lipid translocation across liposome bilayers
[@kubelt_transbilayer_2002;@vehring_flip-flop_2007], can utilize an essentially 
identical data analysis workflow, `flippant` is expected to facilitate analysis
of such experiments as well.
